Human Activity Recognition
==========================

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data 
about personal activity relatively inexpensively. These type of devices are part of the quantified self movement
– a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns
in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular
activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from 
accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts 
correctly and incorrectly in 5 different ways. More information is available from the website 
here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## Objective

Build a predictive model on train data and predict for test data using the caret package functionalities 
neccessisating feature selection due to large number of observations

## Predictive Modelling

Load libaries

```{r}
require(caret)
require(ggplot2)
require(knitr)
require(data.table)
```

Load data in R

```{r}
train_data <- read.csv('pml-training.csv', na.strings=c('', 'NA', 'NULL'))
validation_data <- read.csv('pml-testing.csv', na.strings=c('', 'NA', 'NULL'))
```

## Feature Selection

Remove predictors with NA values

```{r}
na.coldrop <- as.data.frame(which(apply(train_data,2,function(x) {any(is.na(x))})))
train_data <- train_data[,-(na.coldrop[,1])]
validation_data <- validation_data[,-(na.coldrop[,1])]
```

Drop predictors which do not seem to have any relation with output variable 'classe'

* 'X'
* 'user_name'
* 'raw_timestamp_part_1'
* 'raw_timestamp_part_2'
* 'cvtd_timestamp'
* 'new_window'
* 'num_window'

```{r}
train_data <- train_data[,!names(train_data) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
validation_data <- validation_data[,!names(validation_data) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

## Cross Validation

Cross validation was achieved by splitting train_data such that 60% of the train_data was allocated to the 
training and the remainder for testing set

```{r}
sset.seed(138)

inTrain = createDataPartition(train_data$classe, p=0.6, list=FALSE)
training <- train_data[inTrain,]
testing <- train_data[-inTrain,]
dim(training);dim(testing)
dim(training);dim(testing)
```

## Fit Random Forest 

parameters to control the training of the random forest method cross-validation with 4 folds.

```{r}
trControl <- trainControl(method="cv", number=4, verboseIter=T)
modFit.rf <- train(classe ~., data=training, method="rf", trControl=trControl)
```

## Out of Sample Accuracy

```{r}
predout.rf <- predict(modFit.rf, newdata=testing)
error_out <- sum(predout.rf != testing$classe) * 100 / nrow(testing)
error_out
```

## Predictions

```{r}
predictions <- predict(modFit.rf, newdata=validation_data)
predictions
```

function for writing predictions files 

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```

save predictions as txt files indexed by problem_id 

```{r}
pml_write_files(predictions)
```
